---
title: "Making Labelled Volcano Plots From Mass Spectrometry
 Data"
author: "Anthony Doherty"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load the necessary libraries

```{r, libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggplot2)
library(pheatmap)
library(RColorBrewer)
library(ggrepel)
```

### Fix the below path to the location of the files

```{r, setwd}
setwd("/Users/adohert6/Library/CloudStorage/OneDrive-UniversityofEdinburgh/SS18-SSX IP_MS/Results")
```

### Read in Twenty_Four Hour DF & look at column names

```{r, read in data}
Twenty_Four_Hour_DF <-read.table("SS18_SSX_24hr-v-IgG_24hr-data.tsv", header = T)
colnames(Twenty_Four_Hour_DF)
```

### Use dplyr to select columns 1 to 8 and columns 9 and 21 to 24

```{r, select columns}
Twenty_Four_Hour_DF <- Twenty_Four_Hour_DF %>% select(2:8, 9, 21:24)
```

### Filter out any rows in the data frame where the Symbol is Keratin

```{r, filter}
Twenty_Four_Hour_DF <- Twenty_Four_Hour_DF %>% filter(!grepl("KRT", Symbol))
```


### Read in a list of BAF complex genes and make a vector of them

```{r, read in BAF complex genes}
baf_complex_genes <- read.table(("Baf_complex_genes.txt"), header = T)
baf_complex_genes_vector <- baf_complex_genes$Gene
```

### Read in a list of cPRC1 complex genes and make a vector of them

```{r, read in cPRC1 complex genes}
cPRC1_complex_genes <- read.table(("cPRC1_genes.txt"), header = T)
cPRC1_complex_genes_vector <- cPRC1_complex_genes$Gene
```

### Read in a list of vPRC1 complex genes and make a vector of them

```{r, read in vPRC1 complex genes}
vPRC1_complex_genes <- read.table(("vPRC1_genes.txt"), header = T)
vPRC1_complex_genes_vector <- vPRC1_complex_genes$Gene
```

### Show the BAF complex vector

```{r, show the gene subsets}
print(baf_complex_genes_vector)
print(cPRC1_complex_genes_vector)
print(vPRC1_complex_genes_vector)
```

### Make a volcano plot with only the BAF complex genes labelled

```{r, EnhancedVolcano with BAF complex genes}
volcano_plot_Twenty_Four_Hour_Baf <- ggplot(Twenty_Four_Hour_DF, aes(x = logFC, y = NegLogPvalMod)) +
  # Add reference lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50", size = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", size = 0.7) +
  
  # Customize points
  geom_point(aes(color = categ_Mod), 
             size = ifelse(Twenty_Four_Hour_DF$categ_Mod == "Significant", 1.5, 1.5),
             alpha = ifelse(Twenty_Four_Hour_DF$categ_Mod == "Significant", 1, 0.75),
             shape = 16) +  # Simplify shape
  
  # Improved color palette
  scale_color_manual(values = c("Not Significant" = "#6BAED6", "Significant" = "#D7191C")) +
  
  # Scale and axis customization
  scale_x_continuous(
    limits = c(-15, 15),
    breaks = seq(-12, 12, 2)  # Evenly spaced ticks
  ) +
  ylim(c(0, 8)) +
  
  # Add labels for selected genes
  geom_label_repel(aes(label = ifelse(Symbol %in% baf_complex_genes_vector, Symbol, "")),
                   size = 3, box.padding = 0.5, force = 10, 
                   fill = "white", color = "black", 
                   max.overlaps = 500, segment.color = "gray70", segment.size = 0.5) +
  
  # Titles and labels
  labs(
    title = "Volcano Plot: SS18-SSX vs IgG (24-Hour)",
    subtitle = "Highlighting Significant Genes in the BAF Complex",
    x = expression(Log[2]~"Fold Change"),
    y = expression("-log"[10]~"p-value"),
    color = "Protein Status"
  ) +
  
  # Enhanced theme
  theme_minimal(base_size = 14) +  # Clean, minimalistic base theme
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16, color = "darkblue"),
    plot.subtitle = element_text(hjust = 0.5, size = 12, color = "darkgray"),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(size = 12, color = "black"),
    legend.position = "top",  # Position legend at the top
    legend.title = element_text(face = "bold"),
    panel.grid.major = element_line(color = "gray90", linetype = "dashed"),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 1.2)
  )

```

# Print the plot

```{r, Print the plot BAF, fig.width=5, fig.height=8}
print(volcano_plot_Twenty_Four_Hour_Baf)
```

```{r, save plot BAF, warning=FALSE, echo=FALSE}
#ggsave("volcano_plot_Twenty_Four_Hour_vs_Control_BAF.pdf", volcano_plot_Twenty_Four_Hour_Baf, width = 6, height = 6)
```

```{r, EnhancedVolcano with cPRC1 complex genes}
volcano_plot_Twenty_Four_Hour_cPRC1 <- ggplot(Twenty_Four_Hour_DF, aes(x = logFC, y = NegLogPvalMod)) +
  # Add reference lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50", size = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", size = 0.7) +
  
  # Customize points
  geom_point(aes(color = categ_Mod), 
             size = ifelse(Twenty_Four_Hour_DF$categ_Mod == "Significant", 1.5, 1.5),
             alpha = ifelse(Twenty_Four_Hour_DF$categ_Mod == "Significant", 1, 0.75),
             shape = 16) +  # Simplify shape
  
  # Improved color palette
  scale_color_manual(values = c("Not Significant" = "#6BAED6", "Significant" = "#D7191C")) +
  
  # Scale and axis customization
  scale_x_continuous(
    limits = c(-15, 15),
    breaks = seq(-12, 12, 2)  # Evenly spaced ticks
  ) +
  ylim(c(0, 8)) +
  
  # Add labels for selected genes
  geom_label_repel(aes(label = ifelse(Symbol %in% cPRC1_complex_genes_vector, Symbol, "")),
                   size = 3, box.padding = 0.5, force = 10, 
                   fill = "white", color = "black", 
                   max.overlaps = 500, segment.color = "gray70", segment.size = 0.5) +
  
  # Titles and labels
  labs(
    title = "SS18-SSX vs IgG (24-Hour)",
    subtitle = "Highlighting Significant Genes in the cPRC1 Complex",
    x = expression(Log[2]~"Fold Change"),
    y = expression("-log"[10]~"p-value"),
    color = "Protein Status"
  ) +
  
  # Enhanced theme
  theme_minimal(base_size = 14) +  # Clean, minimalistic base theme
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16, color = "darkblue"),
    plot.subtitle = element_text(hjust = 0.5, size = 12, color = "darkgray"),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(size = 12, color = "black"),
    legend.position = "top",  # Position legend at the top
    legend.title = element_text(face = "bold"),
    panel.grid.major = element_line(color = "gray90", linetype = "dashed"),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 1.2)
  )
```

# Print the plot

```{r, Print the plot cPRC1, fig.width=5, fig.height=8}
print(volcano_plot_Twenty_Four_Hour_cPRC1)
```

```{r, save plot cPRC1, warning=FALSE, echo=FALSE}
#ggsave("volcano_plot_Twenty_Four_Hour_vs_Control_cPRC1.pdf", volcano_plot_Twenty_Four_Hour_cPRC1, width = 6, height = 6)
```

```{r, EnhancedVolcano with vPRC1 complex genes}
volcano_plot_Twenty_Four_Hour_vPRC1 <- ggplot(Twenty_Four_Hour_DF, aes(x = logFC, y = NegLogPvalMod)) +
  # Add reference lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray50", size = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray50", size = 0.7) +
  
  # Customize points
  geom_point(aes(color = categ_Mod), 
             size = ifelse(Twenty_Four_Hour_DF$categ_Mod == "Significant", 1.5, 1.5),
             alpha = ifelse(Twenty_Four_Hour_DF$categ_Mod == "Significant", 1, 0.75),
             shape = 16) +  # Simplify shape
  
  # Improved color palette
  scale_color_manual(values = c("Not Significant" = "#6BAED6", "Significant" = "#D7191C")) +
  
  # Scale and axis customization
  scale_x_continuous(
    limits = c(-15, 15),
    breaks = seq(-12, 12, 2)  # Evenly spaced ticks
  ) +
  ylim(c(0, 8)) +
  
  # Add labels for selected genes
  geom_label_repel(aes(label = ifelse(Symbol %in% vPRC1_complex_genes_vector, Symbol, "")),
                   size = 3, box.padding = 0.5, force = 10, 
                   fill = "white", color = "black", 
                   max.overlaps = 500, segment.color = "gray70", segment.size = 0.5) +
  
  # Titles and labels
  labs(
    title = "SS18-SSX vs IgG (24-Hour)",
    subtitle = "Highlighting Significant Genes in the vPRC1 Complex",
    x = expression(Log[2]~"Fold Change"),
    y = expression("-log"[10]~"p-value"),
    color = "Protein Status"
  ) +
  
  # Enhanced theme
  theme_minimal(base_size = 14) +  # Clean, minimalistic base theme
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", size = 16, color = "darkblue"),
    plot.subtitle = element_text(hjust = 0.5, size = 12, color = "darkgray"),
    axis.title = element_text(face = "bold"),
    axis.text = element_text(size = 12, color = "black"),
    legend.position = "top",  # Position legend at the top
    legend.title = element_text(face = "bold"),
    panel.grid.major = element_line(color = "gray90", linetype = "dashed"),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 1.2)
  )
```

# Print the plot

```{r, Print the plot vPRC1, fig.width=5, fig.height=8}
print(volcano_plot_Twenty_Four_Hour_vPRC1)
```

```{r, save plot vPRC1, warning=FALSE, echo=FALSE}
#ggsave("volcano_plot_Twenty_Four_Hour_vs_Control_vPRC1.pdf", volcano_plot_Twenty_Four_Hour_vPRC1, width = 6, height = 6)
```

## Show the session info

```{r, session info}
sessionInfo()
```
